---
title: "Introduction to R"
author: "Alok K. Shukla"
date: "1/23/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Asking for help in R:

```{r Asking for help}
?plot
help(sum)
?help
```

Example of a script

```{r}
x = 100;
y = x+1
print(y)
```

## Scalar Operations

Basic Math Operations
```{r}
x = 100; y <- 69
69 -> y;
sumxy = x+y;
prodxy = x*y;
diffxy = x-y;
quoxy = x/y
```


Other math operations
```{r}
xsq = x^2; lx = log(x); xroot = sqrt(xsq)
```

Complex Numbers, finding modulus and argument:
```{r}
z = x + 1i*y;
magz = abs(z); argz = Arg(z)
lz = log10(z)
is.complex(z)
```

Check for types
```{r}
xstr = "String"
is.numeric(xstr)
```

Printing
```{r}
x =3.4563
print(x,digits=2)
```

## Vector Operations

```{r}
(xvec = c(4,2,6,8,9))  # concatenation, not a 1D matrix
(yvec = c(1,2,3,4,5))
(yvec2 = (1:5))
(yvec3 = seq(from=1,by=1,length.out = 5))
(zvec = c(xvec,yvec))
length(zvec)
dim(zvec)
```


Accessing elements
```{r}
(x4 = xvec[4])
```

Subsetting

```{r}
(x25 = xvec[c(2,5)])
(x2to5 = xvec[2:5])
(xgt5 = xvec[xvec>5])
```

Mathematical Operations

```{r}
xyadd = xvec + yvec; xyprod = xvec*yvec; c=xvec^2 ; xlog = log(xvec)
xyadd
xyprod
c
xlog
```

Statistical Summary
```{r}
summary(xvec)
(mux = mean(xvec))
(sigmax = sqrt(var(xvec)))
(minx = min(xvec))
```

Properties of vector

```{r}
(n2 = length(xvec2 <- rnorm(4))) # on the fly
names(xvec2) = c('First','Second','Third','Fourth')
print(xvec2)
```

## Matrix operations

Creating a matrix

```{r}
(A = matrix(1:20, nrow = 4, ncol = 5)) # ncol optional
(B = matrix(1:20, nrow = 4, ncol =5, byrow = 'T'))
```

Concatenation, row and column-wise
```{r}
rbind(A,B)
cbind(A,B)
```
Accessing elements, display is always row-wise
```{r}
(p = A[2,3])
(q = B[1:3,2:4])
(r = A[3,])
(s = B[,4])
# Vec
xvec[-3]
A[,-3]
```

n-D Array
```{r}
(B2 = array(c(1:3),c(4,5))) # Data, dim
as.vector(B2) # vectorization, stacks columns
dim(B2)
(C2 = array(seq(1,3,length=12),c(2,3,2)))
dim(C2)
C2[,,1]
```


Element-wise operations
```{r}
(C = A+B)
(D = A*B)
(E = A^2)
(F = log(B)/A)
```

Matrix operations
```{r}
# A %*% A
t(A) %*% A
```

Inverse
```{r}
A = matrix(rnorm(16),nrow=4);
(Ainv = solve(A)) # Not as by hand
(Ainv = qr.solve(A)) # More effecient
(Ainv %*% A)
```



# Enough, save workspace variables
save.image(file='var.RData')
# Clear now and load
load('var.RData')

## Plotting

```{r}
x = rnorm(100)
y = rnorm(100)
```

Scatter plots are default

```{r}
plot(x)
plot(x,y)
```

Other kinds

```{r Different plots}
plot(x,type='l') # Default when TS
is.ts(x)
plot(x,type='p')
plot(x,type='b')
plot(x,type='h')
plot(x,type='s')
```

Annotate the plots
```{r}
plot(x,y,type='l',xlab='x axis',ylab='y axis',main='line plot',col='red',font.main=2)
plot(sort(x),y,type='l',xlab='x axis',ylab='y axis',main='line plot',col='red',font.main=2)
```
### Overlays
```{r Overlays}
kvec = (0:99)
xk = sin(2*pi*0.1*kvec)
plot(kvec,xk) # Cant reverse, first plot()
lines(kvec,xk,col='red',lwd=2)
# plot(kvec,xk,col='red',lwd=2,type='l')
```

```{r More overlays}
kvec = (0:99)
xk1 = sin(2*pi*0.1*kvec)  # Sinusoid
xk2 = 0.02*kvec # Line
plot(kvec,xk1,type='b',pch=19,ylim=c(-1,2.5),xlab='Time',ylab='Signal',col='blue')
lines(kvec,xk2,col='red',lty=2,lwd=2)
legstr = c("Sine","Line")
# Look up how to place legends
```

### Multiple plots
```{r}
library(astsa)
```

```{r}
par(mfrow=c(3,1)) # 3 rows
par(mar=c(3,3,3,3)) # Margins LRTB, # lines
par(oma=c(1,1,1,1)) # Paddings
plot(so2,main="SO2 Series",xlab="Year")
boxplot(so2, main="Box plot")
hist(so2,col="grey")
```

# Fancier, using layout
```{r}
M <- rbind(c(1,1),c(2,3))
layout(M)
layout.show(2)
layout.show(3)
par(mar=c(3,3,3,3)) # Margins LRTB, # lines
par(oma=c(1,1,1,1)) # Paddings
plot(so2,main="SO2 Series",xlab="Year",bty='l')
boxplot(so2, main="Box plot")
hist(so2,col="grey",probability = T)
```
 Same using split.screen
 
 
```{r Using Split screen}
split.screen(c(2,1))
screen(1)
par(mar=c(3,3,3,3))
plot(so2,main="SO2 Series",xlab="Year",bty='l')
screen(2)
split.screen(c(1,2),screen=2)
par(oma=c(1,1,1,1))
boxplot(so2, main="Box plot")
screen(4)
hist(so2,col="grey",probability = T)
```


